### 2023-03-30 Writing/Palaeontology/Reviews/response_to_reviewers.tex

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: PALA-11-19-4635-OA}\\
\bigskip
Dear Dr Benson,\\
\bigskip

We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped significantly improve the manuscript.
Unfortunately, due to circumstances out of our control (pandemic, moving continent and new job), it took us far longer to reply to the reviewers requests than anticipated.
Fortunately, however, we have taken all of their comments on board, and respond to their points in detail below.

For improving clarity in this document, reviewer comments are displayed in blue, our responses in normal text and the changes to the manuscript in italic.
Additionally, we've uploaded two versions of the revised manuscript: both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.

@@

\textcolor{blue}{My example points:}

\begin{itemize}

\item{\textcolor{blue}{The abstract lists "shared phylogenetic history" as a factor that causes character correlation in a list of three factors. Approximately the same list of factors is presented in the introduction (p.5, bulletted), where "evolutionary dependence" is stated in place of "shared phylogenetic history". This seems confusing as the two lists seems to serve the same purpose but have non-overalpping content. Shared evolutionary history is a positive thing (we're trying to infer evolutionary history) whereas "evolutionary dependence" of traits seems like it could be negative in some circumstances (for the body elongation example, it may cause it to infer a clade of long-bodied taxa).}}

This is a very helpful point. We have adopted terminological consistency between the abstract and the introduction. We have now standardised the language to "evolutionary dependence" throughout the text and abstract.

@@

\end{itemize}

\section{Referee 1}

\subsection{General notes}
\begin{itemize}
\item{\textcolor{blue}{Please supply your main text in Word format rather than PDF; please supply all figures as separate files and include a list of figure captions at the end of the manuscript rather than embedded in the text. Please check journal submission guidelines:}}

This manuscript was written in \LaTeX\ and no proprietary softwares (e.g. Word) where used. We have added an .odt format version of the manuscript as well as all the figures as separated files with the figure captions at the end of the manuscript.

@@

\begin{itemize}
\item{\textcolor{blue}{Thank you for uploading your data to GitHub. Please replace the link in the text with a citation to a data reference: Guillereme, T., Pastore, A. and Brazeau, M. 20xx. Influence of different modes of morphological character correlation on phylogenetic tree inference. GitHub. https://github.com/TGuillerme/CharactersCorrelation}}

We have now added a proper reference for the data and code.

\item{\textcolor{blue}{We no longer upload any supplementary information on our publisher's website. This information, including figures, needs to be properly curated. Can your supplementary figures be uploaded to GitHub or Figshare? If not, I suggest Dryad (the Palaeontological Association will pay the fee for a Dryad record for any data package associated with a paper in one of its journals). Please contact me directly for an upload link if you wish to use Dryad (other repositories such as MorphoBank are also acceptable). Any text should be supplied in a non-proprietary format (txt files can be supplied alongside a styled pdf version if you prefer). Dryad encourages uploading data in multiple formats if styling is important.}}

We have now added all the supplementary documents as a FigShare item with it's own DOI.

@@

\item{\textcolor{blue}{As above, any data citations in the text should be linked to a data reference including the doi for the record.}}
\end{itemize}

\subsection{References}
\begin{itemize}
\item{\textcolor{blue}{Where you have two references with identical citations (e.g. Guillerme \& Cooper 2016) please list these in the order in which they are cited in the text, irrespective of publication order, or alphabetical order of second author. Therefore Guillerme \& Cooper 2016a should be cited first, whichever reference this is. Having said that, I only found references of Guillerme \& Cooper 2016b so please check this.}}
\end{itemize}

We have now fixed this issue.

\subsection{Figures}

TODO: @@@ I'll do that later

\begin{itemize}
\item{\textcolor{blue}{Please note that although figures may be reproduced in colour in the online edition, there is a fee to reproduce them in print. If a figure appears in colour online and greyscale in print, the same file will be used for both, so please carefully consider your use of colours to make sure that it works in shades of grey.}}
\item{\textcolor{blue}{Please supply all of your figures at a resolution of 600 dpi; preferably in tif format using LZW compression. Embedded photographs are fine at 300 dpi, but if any labelling is included the overall figure will require 600 dpi for printing. Please do not use jpg compression at any stage. Final widths should be either single column (80 mm), 2/3 page width (110 mm) or double column (166 mm). Please see the attached figure
guidelines.}}
\item{\textcolor{blue}{Please confirm the final intended width of each figure (166 mm = full page; 110 mm = 2/3 page; or 80 mm = single column) by adding this to the file name (add 166, 110 or 80).}}
\item{\textcolor{blue}{Fig. 1: Please note that although the journal uses British English, it uses suffix variants - ize, -iza and -izi where appropriate (rather than -ise etc.) This would be changed in the main text but it would be helpful if you could correct any instances on figures (e.g. Randomized, Maximized, Minimized). Similarly, journal style uses single quotes throughout (e.g. 'True' tree)}}
\item{\textcolor{blue}{Fig. 2: Note spelling as above. Please do not refer to colours in the caption if this is intended for printing in colour. Generally, a visual legend is much better on figures anyway (there is room for this in the top right hand corner; it does not require a title).}}
\item{\textcolor{blue}{Figs 3, 4: These figures would require colour both online and in print to be clear. If you do not have funding available to pay the colour printing fee (£150 for one figure, plus £50 for each additional one) please state this clearly in your cover letter as the Palaeontological Association is willing to cover the cost of essential colour where alternative funding is not available (at the Editor-in-Chief's discretion). Please view these two figures at full page width (166 mm wide) and check that all text labels are clearly visible. The axis labels in particular are currently too small; generally, text sizes should be in a range equivalent to 6–10 pt Arial (viewed at 100\%).}}
\end{itemize}

\subsection{Tables}

TODO: @@@ I'll do that later

\begin{itemize}
\item{\textcolor{blue}{Please supply tables either separately, or at the end of the manuscript.}}
\item{\textcolor{blue}{Please note that captions would be altered to include a single, short descriptive phrase, with all additional information/abbreviations moved to a footnote.}}
\end{itemize}

@@

We agree with these two broad categories of correlation types, but we didn’t find them adequate to characterise the range of possible correlations that could be encountered.
Given the nature of our work, we thought ‘atomizing’ the types of correlations possible would be useful in preventing misinterpretations of our results.
Nevertheless, we think the reviewer has raised a really important point that we do not wish to dismiss.
We have modified our discussion of the types of correlation and made specific reference to the two types mentioned by the referee.

@@

We are grateful for this insight and have added it to the manuscript:

\textit{Evolutionary dependence itself requires the accurate resolution of a phylogenetic tree (but see useful methods from Pagel 1994 and Billet and Bardin 2019), and is best determined by independent character sets. The later being commonly hard to get because of the nature of the data: some common type of data (e.g. fossil data) offer very few options of independent character sets and no biological datasets can really be independent. Line @@@.}

\item{\textcolor{blue}{Lines 123 – 124: In order to measure this, we used the following scaled Hamming distance between two characters:
\\
I think that it would be worth comparing this to pairwise dissimilarity among taxa. After all, there have been numerous papers published in Palaeontology that use pairwise dissimilarity (also called phenetic dissimilarity or phenetic distance) to measure disparity. The difference is basically R-mode vs. Q-mode: for phenetic dissimilarity, we count the number of differences between two taxa and divide by the total number of comparable characters to get the proportion of different characters. Here, we count the number of differences between two characters and divide by n-1 to get (almost) the proportion of differing taxa. It basically is like generating scatterplots for every pair of characters, and then counting the number of taxa that fall above or below the line bisecting the axes (0,0 : m,m).}}

In this sense, this metric is used to compare \textit{characters} with each other to create matrices with more or less induced correlation. Our character correlation metric (character difference) is used as an optimisable variable to increase or decrease correlation in our simulated matrices. We believe that the reviewer here suggests comparing character distances (phenetic distance) between \textit{taxa} (cf. between characters). We are not sure what this would bring to the present manuscript since this was not the focal point of this study. In fact, one could argue that this phenetic distance is somehow recovered when inferring the topologies based on the modified and un-modified matrices. But again, this is not what we where trying to explore here.

@@

\textit{we measured both the Robinson-Foulds distance (the differences in clade compositions; Robinson \& Foulds 1981) and the triplets distance (the differences in tips positions; Dobson 1975) between the trees inferred}. Lines @@@

\end{itemize}

\section{Referee 3}

\textcolor{blue}{Sadly however, I am skeptical that this paper really addresses the central issue with character correlation and inferring phylogenies. Particularly, their main conclusion is that character correlation is a 'good' thing, i.e. having increased character correlation actually results in more accurate topologies.}

This isn't quite a correct interpretation of the paper. We appreciate therefore that our explanation may not have been entirely clear and must have caused some confusion. We think, however, it would be an unreasonable standard to expect one paper to solve a monumental problem such as character correlation effects on phylogeny. We don't propose to solve problem in its totality. We have now revised the introduction and discussion to clarify the message of the paper; if character correlation is highly stochastic within a dataset, the individual correlations effectively cancel each other out. Whereas, if correlation is intentionally decreased by the investigator, this can seriously bias the results. Similarly, if it is maximized by the investigator, as expected, it will also bias the result.

@@

\textcolor{blue}{Character correlation should only be a problem for inferring an accurate topology if the correlated characters contain misleading phylogenetic signal and change multiple times in concert on the tree, i.e. show coordinated homoplasy.  This is effectively the 'Intra-organismal dependence' type of character correlation that the authors identify in the introduction, the type of character correlation that is probably of most concern regarding inferring phylogenetic trees. This type of dependent character state change is not only an issue for morphological phylogenetics, but is likely also a problem for molecular data. For example the presence of particular amino acids at particular sites in proteins is fundamental for determining 3D protein structure, limiting the substitutions that can occur at particular sites if proteins are to maintain this 3D structure, which is similar to the tooth occlusion example that is cited in the text. This a particularly relevant and interesting example, especially given that an empirical analysis investigating the performance of dental data from mammals found that they recover phylogenies that are more in conflict with those inferred from molecular data (Sansom et al. 2017) citing character non-independence as a possible factor driving this incongruence. This paper is absent from the bibliography and should certainly be discussed, especially as the conclusions of this paper differ so radically.}

We thank the reviewer for drawing our attention to this paper (Sansom et al. 2017) and the DNA example and have now added that to our manuscript:

\textit{It is also worth noting that the evolutionary and intra-organismal dependences induced correlations are by no means restricted to the use of discrete morphological data. In fact, by essence, these correlations appear to any type of data whether continuous (e.g. linear measurements or morphometric geometric measurements) or non-morphological (e.g. behaviour, or - much more commonly used - molecular data, Sansom et al. 2017.} Lines @@@

@@

\textcolor{blue}{In the discussion (P6. Line 339) the authors state '(1) evolutionary correlation is implied by simulating the characters using Birth-Death trees; and (2) intra-organismal correlation is present in the matrices for the characters randomly simulated but sharing similar evolutionary simulation regimes.  However, the effect of these sources of correlation was out of the scope of this study and would have required a posteriori changes to the matrices'. This means that characters that are correlated due to shared evolutionary history and characters that show coordinated homoplasy are not distinguished by downstream analyses, limiting the applicability of the results for understanding issues raised by real world examples (such as the dental data analysis by Sansom et al. 2017, Sys Bio).}

We do agree that these are valid limitations from this study and have acknowledged them more clearly on lines @@@ as highlighted above.

% This is the text it should refer to: \textit{Third, by not focusing on modeling the sources of character correlation specifically (see section “Distinction between different character correlations” above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see Sansom et al. 2017 for such an application on dental morphological data). However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e. time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.} Lines @@@

@@

\textcolor{blue}{In the conclusions the authors state: 'However, in empirical datasets, if character difference is not driven by selection (e.g. pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.' I am not entirely sure what the authors mean by this sentence. I think that what they are suggesting is that character correlations that arise due to selection will have different impacts on accuracy of phylogenetic reconstructions vs those that arise from drift, which really is what this paper should be trying to address. Instead the analyses pool these types of character correlation rather than trying to disentangle them.}

We have removed this sentence altogether since it was not impactful for the last part of the conclusion. We agree though that it would have been interested to look at how correlation arise between traits in organisms (e.g. through drift or/and selection) and how that would translate into discrete morphological characters but that was outside of the scope of this project.
